---
title: "Valence_PM_thoughts"
author: "Francesco Pupillo"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: true
  warnings: false
  message: false
project:
  type: website
  output-dir: docs
---

## Valence-PM-thoughts project
Script of the analyses 
"Intrusive Thoughts Mediate the Links Between Affective Valence and Prospective Memory"

```{r}
#| message: false
#| warning: false

rm(list= ls())
source("R/0.get_packages.R")

```

```{r}
# change some parameters for plotting
params<-
  theme(
    plot.title = element_text(size = 25),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    axis.text=element_text(size=16),
    legend.text=element_text(size=rel(2)),
    legend.title = element_text(size=rel(2)), 
    strip.text.x = element_text(size=15)
  )

# use colour-blind friendly palette
okabe_ito <- c("YA" = "#0072B2",  # blue
               "OA" = "#D55E00")  # vermillion


# use the labeller
lab_valence <- ggplot2::as_labeller(c(
  post_induction = "Post-induction Change",
  post_task      = "Post-task Change"
))
```


```{r}
#| include: false
# extract valence and PM from raw data\
#| message: false
#| warning: false
#| 
# get the helper function to extract the rt
source("helper_functions/extract_rt.R")

# now source the script that extract the data
source("R/1.extract_files.R")

```

## 
Five participants who did not belong to the age range of both YA and OA were excluded
```{r}
#------------------------------------------------------------------------------#
# Import datasets and exclude participants
#------------------------------------------------------------------------------#
# import the spss dataset
df_wide<-read.csv("group_data/df_wide.csv")
#df_long<-read.csv("group_data/df_long.csv")

# get the spss file
spss_data<-read_spss("PM final data.sav"  )
#------------------------------------------------------------------------------#

#------------------------------------------------------------------------------#
# create age variable
df_wide$agegroup<-ifelse(df_wide$agegroup==1, "YA", "OA")
#df_long$agegroup<-ifelse(df_long$agegroup==1, "YA", "OA")
spss_data$AgeGroup<-ifelse(spss_data$AgeGroup==1, "YA", "OA")

single_part<- df_wide %>%
  group_by(participant) %>%
  slice(1)

print(paste("Total number of participants is", nrow(single_part)))

# plot the distribution of age as a function of agegroup
spss_data$AgeGroup<-factor(spss_data$AgeGroup, levels = c("YA", "OA"))

# exclude participants who do not belong to the classical age range
age_excl<-spss_data$Nb[(spss_data$Age>35 & spss_data$AgeGroup=="YA") |
                        (spss_data$Age<60 & spss_data$AgeGroup=="OA") ]

# exclude those participants 
spss_data<-spss_data[!spss_data$Nb %in% age_excl,]
df_wide<-df_wide[!df_wide$participant %in% age_excl,]
single_part<-single_part[!single_part$participant %in% age_excl,]
#df_long<-df_long[!df_long$participant %in% age_excl,]


age_plot<-ggplot(spss_data, aes(x = Age))+
  geom_histogram()+
  facet_wrap(.~AgeGroup)+
  theme_classic()

print(age_plot)

# check the number of participants
kable(df_wide %>%
        group_by(participant) %>%
        slice(1) %>%
        group_by( agegroup)%>%
        tally())

#--------------------

```



## Sociodemographics
First, apply exclusion for participants who do not belong to the age range
and get demographics
```{r}
source("R/2.sociodemographics.R")

kable(gender_part, caption = "Gender by Agegroup")

kable(part_demo, caption = "Sociodemographics")

```

## Sociodemographics
Compare the sociodemographics between young and old
```{r}
source("R/3.compare_sociodemo.R")

summary(mod_health)

summary(mod_educat)

summary(mill_hill)

summary(digit_symbol)

# create a contingency table to check whther the distribution of gender is equal
contingency_table_gender<-table(single_part[, c("agegroup", "gender")])
chi_square_test <- chisq.test(contingency_table_gender)
chi_square_test

```

Valence reshaped
```{r}
# now call  the script to the reshape the valence
source("R/4.valence_reshape.R")

# participant check
# check the number of participants
kable(long_df_valence %>%
        group_by(participant) %>%
        slice(1) %>%
        group_by( agegroup)%>%
        tally())

kable(long_df_valence_diff %>%
        group_by(participant) %>%
        slice(1) %>%
        group_by( agegroup)%>%
        tally())


```

Preprocessing of valence. 
Derivation of post-induction and post valence measures and visualization
```{r}
source("R/5.valence_preproc.R")
```
## Valence analyses
```{r}
source("R/6.valence_analysis.R")
```


## PM analysis
```{r}
source("helper_functions/logit_transform.R")


source("R/7.PM_main_analysis.R")

```

## Intrusive thoughts

```{r}
source("R/8.intrusive_thoughts.R")
```

## Mediation analyses
```{r}
#| warning: false
source("R/9.mediations.R")
```



