---
title: "Valence_PM_thoughts"
author: "Francesco Pupillo"
format:
  html:
    toc: true
    code-fold: true
    output-file: index.html
execute:
  echo: true
  warnings: false
  message: false
---

## Valence-PM-thoughts project
Script of the analyses 
"Intrusive Thoughts Mediate the Links Between Affective Valence and Prospective Memory"

```{r}
#| message: false
#| warning: false

rm(list= ls())
source("R/0.get_packages.R")

```

```{r}
#| message: false
#| warning: false
# change some parameters for plotting
params<-
  theme(
    plot.title = element_text(size = 35),
    axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28),
    axis.text=element_text(size=26),
    legend.text=element_text(size=rel(4)),
    legend.title = element_text(size=rel(4)), 
    strip.text.x = element_text(size=25)
  )

# use colour-blind friendly palette
okabe_ito <- c("YA" = "#0072B2",  # blue
               "OA" = "#D55E00")  # vermillion


# use the labeller
lab_valence <- ggplot2::as_labeller(c(
  post_induction = "Post-induction Change",
  post_task      = "Post-task Change"
))
```


```{r}
#| include: false
#| message: false
#| warning: false



# get the helper function to extract the rt
source("helper_functions/extract_rt.R")

# now source the script that extract the data
source("R/1.extract_files.R")

```

## 
Five participants who did not belong to the age range of both YA and OA were excluded
```{r}
#| message: false
#| warning: false
#------------------------------------------------------------------------------#
# Import datasets and exclude participants
#------------------------------------------------------------------------------#
# import the spss dataset
df_wide<-read.csv("group_data/df_wide.csv")

# now we need to add the order of the conditions, i.e. which condition is presented
# first, which one second, and which one third
# there are six conditions, repeated within participants.
# they can be found in the files "Age, Mood and PM Project Participant Code.xlsx
# They are the following

conditions <- matrix(
  c(
    "Positive", "Negative", "Neutral",
    "Positive", "Neutral",  "Negative",
    "Negative", "Positive", "Neutral",
    "Negative", "Neutral",  "Positive",
    "Neutral",  "Negative", "Positive",
    "Neutral",  "Positive", "Negative"
  ),
  
  nrow = 6,
  byrow = TRUE
)

# I need to repeat it 11 times to get all participants
cond_all<-conditions[rep(1:nrow(conditions), 11), ]

# get the order variable
df_wide$order<-NA
for (n in 1:nrow(df_wide)){ # loop through the observations (conditions)
  
  if (df_wide$participant[n]<100){ # if participant number was not modified
    
    c_order<- cond_all[df_wide$participant[n],]
    
    # in which position is that condition for that participant?
    c_num<-which(c_order == df_wide$cond[n])
    
    
  } else { # if we added 100 to the participants (older adults) to distinguish
    # them from the YA
    
    c_order<- cond_all[df_wide$participant[n]-100,]
    
    c_num<-which(c_order == df_wide$cond[n])
    
  }
  
  df_wide$order[n]<-c_num
  
  
  
}


# get the spss file
spss_data<-read.csv("PM final data.csv"  )
#------------------------------------------------------------------------------#

#------------------------------------------------------------------------------#
# create age variable
df_wide$agegroup<-ifelse(df_wide$agegroup==1, "YA", "OA")
#df_long$agegroup<-ifelse(df_long$agegroup==1, "YA", "OA")
spss_data$AgeGroup<-ifelse(spss_data$AgeGroup==1, "YA", "OA")

single_part<- df_wide %>%
  group_by(participant) %>%
  slice(1)

print(paste("Total number of participants is", nrow(single_part)))

# plot the distribution of age as a function of agegroup
spss_data$AgeGroup<-factor(spss_data$AgeGroup, levels = c("YA", "OA"))

# exclude participants who do not belong to the classical age range
age_excl<-spss_data$Nb[(spss_data$Age>35 & spss_data$AgeGroup=="YA") |
                        (spss_data$Age<60 & spss_data$AgeGroup=="OA") ]

# exclude those participants 
spss_data<-spss_data[!spss_data$Nb %in% age_excl,]
df_wide<-df_wide[!df_wide$participant %in% age_excl,]
single_part<-single_part[!single_part$participant %in% age_excl,]
#df_long<-df_long[!df_long$participant %in% age_excl,]


age_plot<-ggplot(spss_data, aes(x = Age))+
  geom_histogram()+
  facet_wrap(.~AgeGroup)+
  theme_classic()

print(age_plot)

# check the number of participants
kable(df_wide %>%
        group_by(participant) %>%
        slice(1) %>%
        group_by( agegroup)%>%
        tally())

#--------------------

```



## Sociodemographics
First, apply exclusion for participants who do not belong to the age range
and get demographics
```{r}
source("R/2.sociodemographics.R")

kable(gender_part, caption = "Gender by Agegroup")

kable(part_demo, caption = "Sociodemographics")

```

## Sociodemographics
Compare the sociodemographics between young and old
```{r}
#| message: false
#| warning: false
source("R/3.compare_sociodemo.R")

summary(mod_health)

summary(mod_educat)

summary(mill_hill)

summary(digit_symbol)

# create a contingency table to check whther the distribution of gender is equal
contingency_table_gender<-table(single_part[, c("agegroup", "gender")])
chi_square_test <- chisq.test(contingency_table_gender)
chi_square_test

```

Valence reshaped
```{r}
#| message: false
#| warning: false
# now call  the script to the reshape the valence
source("R/4.valence_reshape.R")

# participant check
# check the number of participants
kable(long_df_valence %>%
        group_by(participant) %>%
        slice(1) %>%
        group_by( agegroup)%>%
        tally())

kable(long_df_valence_diff %>%
        group_by(participant) %>%
        slice(1) %>%
        group_by( agegroup)%>%
        tally())


```

Preprocessing of valence. 
Derivation of post-induction and post valence measures and visualization
```{r}
#| message: false
#| warning: false
source("R/5.valence_preproc.R")
```
## Valence analyses
```{r}
#| message: false
#| warning: false
source("R/6.valence_analysis.R")
```

## additional arousal analysis
```{r}
#| message: false
#| warning: false
#| include: false
source("R/6S.arousal_analysis.R")

```


## PM analysis
```{r}
#| message: false
#| warning: false
source("helper_functions/logit_transform.R")


source("R/7.PM_main_analysis.R")

# check if the order (repetition) of the task affected PM
long_df_merged$order<-as.numeric(long_df_merged$order)

# center order
long_df_merged$order_c<-ifelse(long_df_merged$order==1, -1, 
                               ifelse(long_df_merged$order==2, 0, 
                                      1))

long_df_merged$DS.c<-scale(long_df_merged$DS)

long_df_merged$Mill_Hill.c<-scale(long_df_merged$Mill_Hill)


reg.s_order<-lmer(PM_task_lenient_av_log~order_c*agegroup.c+DS.c+Mill_Hill.c+aft_PM_minus_after_ind.s*agegroup.c + 
              valence_aftIND_min_bef.s*agegroup.c+(1+order_c|participant), 
            data = long_df_merged[ (is.na(long_df_merged$MOCA)|long_df_merged$MOCA>=26),])
summary(reg.s_order)

tab_model(reg.s_order, file = "main_regression_table.html")


anova(reg.s_order)


vif(reg.s_order)

eta_squared(reg.s_order, partial = T, alternative = "two.sided")

# check the interaction between age and DS
age_DS<-lmer(PM_task_lenient_av_log~agegroup*DS.c+(1|participant), 
            data = long_df_merged[ (is.na(long_df_merged$MOCA)|long_df_merged$MOCA>=26),])

summary(age_DS)


interact_plot(age_DS,
           
              pred = DS.c,
      
              modx = agegroup,
       
              interval = TRUE,
              legend.main = "Age group"
              )+
  labs(x = "Fluid intelligence (centered)", 
       y = "PM performance")+
  theme_classic()+
  labs(colour = "Age group")

# without centering
reg_order<-lmer(PM_task_lenient_av_log~ +DS+order_c+aft_PM_minus_after_ind*agegroup.c + 
            valence_aftIND_min_bef*agegroup.c+(1|participant), 
          data = long_df_merged[ (is.na(long_df_merged$MOCA)|long_df_merged$MOCA>=26),])


Anova(reg_order, type =2)



# 1) Predicted curve for post-induction (hold post-task at 0)
pred_T2T1 <- ggpredict(
  reg_order,
  terms = c("valence_aftIND_min_bef", "agegroup.c"),
  condition = c(aft_PM_minus_after_ind.s = 0)   # set the other predictor to 0 (its natural reference)
) %>%
  as_tibble() %>%
  mutate(measure = "post_induction",
         x = x)   # keep x for clarity

# 2) Predicted curve for post-task (hold post-induction at 0)
pred_T3T2 <- ggpredict(
  reg_order,
  terms = c("aft_PM_minus_after_ind", "agegroup.c"),
  condition = c(valence_aftIND_min_bef = 0)
) %>%
  as_tibble() %>%
  mutate(measure = "post_task",
         x = x)

# 3) Combine and plot
pred_all <- bind_rows(pred_T2T1, pred_T3T2)

print(
ggplot(pred_all, aes(x = x, y = predicted, colour = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group),
              alpha = 0.15, colour = NA) +
  geom_line(size = 1) +
  facet_wrap(~ measure, nrow = 1, labeller = lab_valence) +
  labs(x = "Valence", y = "Predicted PM accuracy", colour = "Age group", fill = "Age group") +
  scale_color_manual(values = c("-0.5" = "#0072B2",  # blue
                                "0.5" = "#D55E00"),  # vermillion, 
                     labels = c("-0.5" = "YA", "0.5" = "OA")) +
  #scale_fill_manual(values = c(YA = "darkorange", OA = "darkgreen")) +  # CI ribbons
  scale_fill_manual(values = c("-0.5" = "#0072B2",  # blue
                               "0.5" = "#D55E00"),  # vermillion, 
                    labels = c("-0.5" = "YA", "0.5" = "OA"))  +  # CI ribbons
  geom_vline(xintercept = 0)+
  theme_classic()+
  theme(legend.position = "none")+
  params

)
```

## Intrusive thoughts

```{r}
#| message: false
#| warning: false
source("R/8.intrusive_thoughts.R")
```

## Mediation analyses
```{r}
#| warning: false
source("R/9.mediations.R")
```

## Create a correlation table with all the variables
```{r}
df_for_corr <- long_df_merged %>%

group_by(participant) %>%
summarise(
agegroup = first(agegroup),
education_y = first(education_y),
health = first(health),
fluid_intelligence = first(DS), 
crystallized_intelligence = first(Mill_Hill),
general_intrusive_thoughts = first(tcaq_g),
task_related_intrusive_thoughts = mean(c_tcaq, na.rm = T),
PM_performance = mean(PM_task_lenient_av, na.rm = T),
#baseline_valence = first(baseline_valence),
post_induction_valence_change = mean(valence_aftIND_min_bef, na.rm = TRUE),
post_task_valence_change = mean(aft_PM_minus_after_ind, na.rm = TRUE),
)

str(df_for_corr)

# confert age group to numeric
df_for_corr$agegroup<-ifelse(df_for_corr$agegroup=="YA", -0.5, 0.5)
cor_tbl<-corr.test(df_for_corr[, 2:ncol(df_for_corr)], use = "pairwise", method = "pearson", adjust = "none")

corr_table <- round(cor_tbl$r, 2)
p_table <- round(cor_tbl$p, 4)

sink("correlation_table.txt")
cat("Correlation matrix\n")
print(round(cor_tbl$r, 2))

cat("\nP-value matrix\n")
print(round(cor_tbl$p, 4))
sink()

```


